Weather and fuel as modulators of grassland fire behavior in the northern Great Plains

### 2023-10-12 paper/springer/CoverLetter/CoverLetter_R2.tex
surface update

\vspace{-5em}

@@

Please consider the attached submission for publication in \journal.

This submission is motivated by the positive experience my colleague and I recently enjoyed publishing in \journal~(DOI \href{https://doi.org/10.1007/s00267-022-01659-y}{10.1007/s00267-022-01659-y}). 
I have decided to move on with this work after a lack of response from the Editor in Chief at the journal to which we initially submitted, following mishandling by an Associate Editor. 
After our initial review, the paper was understandably rejected and a resubmission requested, due to the extent of the reviewer comments. 
Each were productive, and we responded completely (these responses are attached). 
However, despite no reviewer criticism on the resubmitted manuscript, the (presumably new?) AE assigned to the resubmission rejected the manuscript on the grounds that our approach to measuring two-dimensional rate of spread violated assumptions that are in fact limited to one-dimensional rate of spread. 
But recently, Dr. Mark Finney\textemdash almost certainly the leading wildland fire scientist in the US, if not the world\textemdash led a textbook in which his team described the superiority of 2-D measurements over 1-D measurements. 
We brought this to the EiC's attention in our appeal of the AE decision, but have gone several weeks without a response.

As these data are of great interest to our stakeholders, we need to press on with publication. 
I think the scope and audience of \journal~is a good fit for us. 
We have explicitly referenced Finney et al's textbook into our Methods section to make this point clear. 
While I understand \journal~has no formal Portable Peer-Review process, I hope the editorial staff might take this context into account and consider an expedited review process.

>>>>

%\vspace{-5em}

@@

We appreciate the opportunity to revise our submission for publication in \journal.
We've taken the editor's comments about further review seriously. 
We trust the editor will find that each of the second reviewer's comments have been fully replied to in this second round of review. 
Unfortunately the new major revisions period of 14 days is prohibitively short, and our attention in the previous review was focused on the detailed critique of the first reviewer. 
In this latest round, we have attended to all of the second reviewer's points and appreciate the additional detail the reviewer provided.

### 2023-10-12 paper/springer/GrasslandFireBehaviorNGP_R2.tex
surface update

Prescribed fire is used widely for ecosystem management around the world \citep{weir2022}. 
More than simply the result of combustion of vegetation, wildland fire behavior is multi-faceted, with different components producing different effects on the surrounding environment and organisms within. 
Most wildland fire scientists describe fire behavior in terms of \emph{rate of spread}---how quickly a flame front moves through a fuelbed---and \emph{intensity}---a suite of measurements of how much energy is released by combustion, often expressed as a rate of energy release over time \citep{mcgranahan2021a}.

@@

\emph{Biotic factors} relate principally to the amount and nature of
plant biomass available for combustion. Overall, energy release rates
increase as more fuel is available to burn. The structure and
arrangement of vegetation is also important. Greater fuel load
attributable to longer time-since-fire increases fire temperature, and
spatial variability in fuel load and patchy distribution of fine fuels
in turn drive variability in fire behavior
\citep{patten1984, gibson1990, gomes2020}. Furthermore, fine-leaved
grasses burn more completely and hotter than an equal mass of forbs
\citep{wragg2018}. Finally, fuel moisture content is an especially
important driver outside of the highly-cured context of wildfire seasons
\citep{sparling1966, kidnie2015}. Together, variability in flammability
traits and curing rates among species that comprise grassland fuelbeds
contributes to variability in fire behavior
\citep{cruz2015, kidnie2015, mcgranahan2016, cardoso2018}.

@@

Despite its popularity among fire ecologists, temperature alone is a
poor response variable fraught by several issues in collecting and
interpreting thermocouple data \citep[see review by][]{mcgranahan2020}.
Firstly, a considerable amount of variability in temperature is
attributable to sensor placement relative to both the ground and the
fire. There is neither a standard for placing thermocouple probes in the
wildland fire environment nor consistency in vertical temperature
profiles. Most observations of surface fire temperature profiles
describe an inverse, linear relationship between height and temperature
\citep{smith1966, patten1984, archibold2003}, although
\citet{ramsay1996} found the highest temperatures at the top of a 1 m
profile and the lowest temperatures at 30 cm, while \citet{frost1987}
and \citet{bailey1980} present evidence that the highest temperatures
occur midway up the profile. At least some of this variability might be
due to differences in surface vs.~flame temperature among head and back
fires \citep{trollope1978}.

Secondly, many factors that contribute to variability in temperatures
recorded by thermocouples are attributable to the nature of the sensor
rather than the nature of the fire \citep[e.g.,][]{walker1968}. Thus,
reporting `device temperatures' alone impedes comparisons between
studies; \citet{bova2008} present a standard calibration of thermocouple
probes, while \citet{mcgranahan2021b} simply uses the timestamps of peak
heating across an array of several thermocouples to calculate rate of
spread. Using rate of spread as the response variable makes moot the
third issue with temperature as a fire behavior metric: temperature of
the media around a probe is a poor proxy for the thermal experience of
an organism. Measures of intensity or energy flux are more biologically
relevant \citep{kremens2012, smith2016}.

In the North American Great Plains, most reports of grassland fire
behavior consist of temperatures derived from thermocouples, and there
are few data on rate of spread. Soil surface temperatures in South
Dakota tallgrass prairie ranged from 200-500\(^\circ\)C during spring
burns, and were greatest under lower fuel loads \citep{ohrtman2015};
fires in Saskatchewan mixed grass prairie exceeded 300\(^\circ\)C 5-10
cm above the soil surface in spring, summer, and fall
\citep{archibold2003}. Mean temperatures in experimental burns in
eastern Montana ranged from 172-222\(^\circ\)C in the summer to 253\(^\circ\)C in the spring \citep{strong2013, russell2015}. To our
knowledge, no studies on grassland fire in the northern Great Plains
region has explicitly tested the effect of fire weather on fire
behavior.

@@

\hypertarget{data-collection}{%
\subsection{Data collection}\label{data-collection}}

@@

At each 1 m triangular microplot---the observational unit in the nested plot design---we used four thermocouples connected to a single FeatherFlame datalogger. 
Three thermocouples measured flame temperature 15 cm above the soil surface at each 1 m vertex while a fourth thermocouple recorded soil surface temperature at a representative point within the 1 m array. 
Beads of the thermocouple probes extended at least 3 cm from the supporting apparatus, to which the insulated lead was attached with wire. 
The soil thermocouple was placed on mineral ground, perpendicular to the soil surface, below plant litter.

@@

\textbf{Fire weather data} were obtained after each fire from records
made available by the North Dakota Agriculture Weather Network, the
statewide mesonet system with sensor arrays at both experimental
stations. We downloaded hourly relative humidity (\%), dew point
(\degC), air temperature (\degC), and average wind speed (m s\textsuperscript{-1}). From these data we calculated atmospheric vapor pressure ($e$) and saturation vapor pressure ($e_s$) and used these
quantities to determine the vapor pressure deficit ($VPD = e - e_s$)
for the hour in which each fire behavior observation occurred. These
data capture hourly trends in weather at the meso-gamma scale \citep[2-20 km;][]{orlanski1975}, and are reliably connected to our fire behavior measurements via time stamps provided by the dataloggers. 
The Hettinger mesonet array is 3--8 km from burned pastures, while the Central Grasslands mesonet array is 1--7.5 km from burned pastures. We
found a high degree of consistency between these meso-scale data and
fire weather records made on the fireline during operational periods,
and the open rangeland physiognomy with flat to rolling terrain
precludes substantial microsite differences in weather between these
records and the conditions at each fire behavior sample point.

\hypertarget{data-analysis}{%
\subsection{Data analysis}\label{data-analysis}}

@@

\textbf{Regression analysis.---}We assessed weather and fuel effects on three fire behavior response variables: Maximum flame temperature (\degC) 15 cm above the soil surface (mean of three thermocouples), maximum soil surface temperature (\degC; single thermocouple), and rate of spread (m min\textsuperscript{-1}) through each 1 m triangular microplot. 
Because all three response variables were best modeled with a gamma distribution, we fit generalized linear mixed-effect regression models for each response with the \texttt{glmer} function from the \emph{lme4} package in \textsf{R} \citep{bates2015}. 
Fixed effects consisted of weather and fuel variables, as described above. 
The
random-effect term was constructed to account for random variance among
locations, spatial non-independence within locations and nested variance
within sample plots, and the effect of repeated measurements within each
location. Due to concerns about collinearity between relative humidity
and vapor pressure deficit because they are derived from the same
variables, vapor pressure deficit was excluded from regression analysis
for all three response variables.

\hypertarget{results}{%
\section{Results}\label{results}}

@@

Most measures of fuel, fire weather, and fire behavior showed
considerable variability within each location, although rates of spread
were generally low (Fig.~\ref{DataSummary}). 
Principal Components Analysis indicated fire behavior patterns were consistent across locations ($P$ = 0.11), although the fall season in Hettinger---our semi-arid location in southwestern North Dakota---tended to have drier air and hotter fires (Fig.~\ref{PCA}). 
Spring fires in the Central Grasslands were conducted under warmer and more evaporative (VPD) conditions than fall fires at
Hettinger. The first two axes of the Principal Components Analysis (Fig.
2) explained 86\% of overall variance in the fire behavior dataset. The
first axis (52\% variance explained) was most strongly associated with
flame temperature and rate of spread, while the second axis was more
strongly associated with soil surface temperature. Dew point was
marginally related ($P~<$ 0.05) and inversely associated with
flame temperatures and rate of spread.

\begin{figure} 
\includegraphics[width = 1\columnwidth]{pca_gg-1.pdf}
\caption{Principal Components Analysis of fire behavior data (response
variables in blue; rate of spread (ROS), temperature above surface
(flame \degC), and temperature at soil surface (soil \degC) for prescribed
burns on rangeland at Hettinger (H), in southwestern North Dakota, and
Central Grasslands (CG), in central North Dakota. No difference between
locations ($P$ = 0.11). 
Total variance explained in these two axes = 86\%.}
\label{PCA}
\end{figure}

Fires at both locations were characterized by considerable variability
among sub-plots. Above the soil surface, half of the fires (13) exceeded
325\(^\circ\)C, but only four of those fires had \textgreater50\% of
individual sample plots within the burns reach an average of
325\(^\circ\)C. Sparse fuels meant that fire did not spread to some
individual plots in some burns despite strip ignitions. Among plots that
burned, fewer than half exceeded 100\(^\circ\)C at the soil surface (49
of 121 plots). A majority of the fires (18) had at least one plot exceed
100\(^\circ\)C at the soil surface, seven had over half of the plots
exceed 100\(^\circ\)C at the soil surface, and only one fire had all
plots reach this (a spring Central Grasslands fire). Only 10 fires
reached or exceeded 325\(^\circ\)C on the soil surface, with most of
these fires reaching this point in only one plot and never in more than
half of the plots.

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

@@

Mean temperatures recorded in this study are consistent with other
reports from northern rangelands. In central Alberta grassland,
\citet{bailey1980} observed that surface temperatures varied between
110\(^\circ\)C and 165\(^\circ\)C for backfires and headfires,
respectively, and headfires averaged 200\(^\circ\)C 15 cm above the
ground; temperatures generally tracked with fuel load. Surface fires
through jack pine barrens in Ontario had a similar range as ours
\citep[140-545\(^\circ\)C,][]{smith1966}. In our study, mean 15-cm
temperatures were 225\(^\circ\)C during spring burns in central North
Dakota and 250\(^\circ\)C during fall burns in southwestern North
Dakota; surface temperatures at both locations generally averaged just
above 100\(^\circ\)C (Fig. 1).

Discrepancies between our data and others from the region are consistent
with what would be expected when differences in the fire environment are
considered. For example, \citet{ohrtman2015} reported a wide range of
maximum temperatures at the soil surface---150-500\(^\circ\)C---that was
generally explained by variability in annual productivity and clipping
frequency, which altered fuel load. Our fires were also cooler than
those reported by \citet{archibold2003} in Saskatchewan: using the
mid-point of observations made at 10 cm and 20 cm as a comparison to our
15-cm values, spring fires reached 314\(^\circ\)C and fall fires reached
298\(^\circ\)C. But \citet{archibold2003} also reported substantially
lower fuel moisture in each season and they had approximately three
times the fuel load, likely due to an absence of grazing on the remnant
prairie. A previous study reported similar results---temperatures
approaching 500\(^\circ\)C when fuel loads averaged 2.8-4.5 t
ha\(^{\text{-}1}\) \citep{archibold1998}. With greater variability in
fuel load and fuel moisture, we might also expect to see these factors
have greater influence on aboveground flame temperatures. For example,
in Colorado, \citet{augustine2014} observed a strong linear relationship
between fuel load and temperatures ranging from 60-200\(^\circ\)C, but
their fuel load also ranged from 0.2 to 1.2 t ha\(^{\text{-}1}\).

Fires at both of our locations spread much more slowly than most reports
from other grasslands, due primarily to high fuelbed moisture content
and little opportunity for mitigation by wind or lower atmospheric
moisture (Fig. 1). Perhaps the most variability in fire spread was
reported by \citet{sneeuwjagt1977} from prescribed grass fires in
California and Washington, where rates of spread ranged from 0.2-61 m
min\(^{\text{-}1}\). From grassland fires in South Africa and Kansas,
USA, \citet{trollope2002} reported average spread rates of 24 and 32 m
min\(^{\text{-}1}\), respectively, for head fires and 0.12 and 0.14 m
min\(^{\text{-}1}\), respectively, for back fires. In a tallgrass
prairie in Texas, \citet{clements2019} recorded fire spreading between
72-150 m min\(^{\text{-}1}\) with the wind and 48 m min\(^{\text{-}1}\)
for flanking fires. Likewise, fires through cured grass fuelbeds in
Australia spread much more rapidly than we observed---up to 18-180 m
min\(^{\text{-}1}\) \citep{cheney1993, cheney1995, cruz2015}. Through
partially-cured stands, though, \citet{cruz2015} found spread rates
dropped to 3-44 m min\(^{\text{-}1}\), approaching our location averages
of 2.1 and 3.1 m min\(^{\text{-}1}\). Consistent with our finding that
only wind speed had a statistically-significant effect on increasing
rate of spread (Fig. 3), \citet{cheney1993} found that wind was by far
the most important variable to spread rate.

@@

Much is made of the difference in fire behavior between head and back
fires in the fire ecology literature, and while we expect these
differences translate to different fire effects in our system, making
distinctions between fire types is difficult in both our data and our
management. \citet{trollope1978} emphasized that while head fires move
faster and generally release more energy, back fires effect greater
heating at the ground level. One would expect, then, that back fires
would have more opportunity to burn down through even thick \emph{P.
pratensis} litter to mineral soil. Unfortunately, our results offer
little insight into what fuel or weather variables enhance litter
combustion, likely because most of our fires never got very hot at the
soil surface---50\% of our observations were less than 60\(^\circ\)C,
and 60\% less than 100\(^\circ\)C (Fig.~\ref{DataSummary}). 
Nor can we differentiate the direction of fire spread with the current trigonometry applied to the
triangular thermocouple arrays \citep{simard1984}, although it would
theoretically be possible to compare spread direction to wind direction
if the latter data were available at a fine enough scale.

>>>>

Prescribed fire is used widely for ecosystem management around the world \citep{weir2022}. 
More than simply the result of combustion of vegetation, wildland fire behavior is multi-faceted, with different components producing different effects on the surrounding environment and organisms within. 
Unlike wildfire events, in which incident commanders are most interested in understanding fire behavior to ensure safe suppression operations, often during extremely dry conditions, prescribed fire managers seek to understand fire behavior to achieve specific vegetation  objectives under a range of weather and fuel moisture conditions. 
In either context, wildland fire scientists describe fire behavior in terms of \emph{rate of spread}---how quickly a flame front moves through a fuelbed---and \emph{intensity}---a suite of measurements of how much energy is released by combustion, often expressed as a rate of energy release over time \citep{mcgranahan2021a}.

@@

\emph{Biotic factors} relate principally to the amount and nature of
plant biomass available for combustion. Overall, energy release rates
increase as more fuel is available to burn. The structure and
arrangement of vegetation is also important. Greater fuel load increases fire temperature, and
spatial variability in fuel load and patchy distribution of fine fuels
in turn drive variability in fire behavior
\citep{patten1984, gibson1990, gomes2020}. Furthermore, fine-leaved
grasses burn more completely and hotter than an equal mass of forbs
\citep{wragg2018}. 
Finally, while highly cured, uniformly low fuel moisture content in active wildfire seasons contributes to greater energy release and faster rates of spread, fire behavior outside of these seasons\textemdash when prescribed fire is typically applied\textemdash is often limited by high fuel moisture content \citep{sparling1966, kidnie2015}.
Together, variability in flammability
traits and curing rates among species that comprise grassland fuelbeds
contributes to variability in fire behavior
\citep{cruz2015, kidnie2015, mcgranahan2016, cardoso2018}.

@@

Despite its popularity among fire ecologists, temperature alone is a
poor response variable fraught by several issues in collecting and
interpreting thermocouple data \citep[see review by][]{mcgranahan2020}.
A considerable amount of variability in temperature is attributable to sensor placement relative to both the ground and the fire (e.g., see conflicting reports of peak temperatures relative to soil surface described in \citet{smith1966, patten1984, archibold2003, ramsay1996, frost1987, bailey1980}). 
Additionally, many factors that contribute to variability in temperatures recorded by thermocouples are attributable to the nature of the sensor rather than the nature of the fire \citep[e.g.,][]{walker1968}, impeding comparisons between studies \citep{bova2008}.
And because temperature of the media around a probe is a poor proxy for the thermal experience of an organism, measures of intensity or energy flux are more biologically relevant \citep{kremens2012, smith2016}.
Rate of spread, rather than temperature, is a factor in calculating fire intensity \citep{byram1959}, and can be derived from  thermocouple arrays by simply using the timestamps of peak heating to calculate rate of spread \citep{mcgranahan2021b, finney2021}.

In the North American Great Plains, most reports of grassland fire behavior consist of temperatures derived from thermocouples, and there are few data on rate of spread. 
Soil surface temperatures in South Dakota tallgrass prairie ranged from 200-500\(^\circ\)C during spring burns, and were greatest under lower fuel loads \citep{ohrtman2015}; fires in Saskatchewan mixed grass prairie exceeded 300\(^\circ\)C 5-10 cm above the soil surface in spring, summer, and fall \citep{archibold2003}. 
Mean temperatures in experimental burns in eastern Montana ranged from 172-222\(^\circ\)C in the summer to 253\(^\circ\)C in the spring \citep{strong2013, russell2015}. 
To our knowledge, no studies on grassland fire in the northern Great Plains has explicitly tested the effects of fuel moisture and fuel load, or fire weather, on fire behavior.

@@

\subsection{Data collection}

@@

At each 1 m triangular microplot---the observational unit in the nested plot design---we used four thermocouples connected to a single FeatherFlame datalogger. 
Three thermocouples measured flame temperature 15 cm above the soil surface at each 1 m vertex while a fourth thermocouple recorded soil surface temperature at a representative point within the 1 m array. 
We chose 15 cm for our single observation as the midpoint of 10 and 20 cm observations reported elsewhere \citep[e.g.,][]{archibold2003}.
Beads of the thermocouple probes extended at least 3 cm from the supporting apparatus, to which the insulated lead was attached with wire. 
The soil thermocouple was placed on mineral ground, perpendicular to the soil surface, below plant litter.

@@

\textbf{Fire weather data} were obtained after each fire from records
made available by the North Dakota Agriculture Weather Network, the
statewide mesonet system with sensor arrays at both experimental
stations. We downloaded hourly relative humidity (\%), dew point
(\degC), air temperature (\degC), and average wind speed (m s\textsuperscript{-1}). From these data we calculated atmospheric vapor pressure ($e$) and saturation vapor pressure ($e_s$) and used these
quantities to determine the vapor pressure deficit ($VPD = e - e_s$)
for the hour in which each fire behavior observation occurred. These
data capture hourly trends in weather at the meso-gamma scale \citep[2-20 km;][]{orlanski1975}, and are reliably connected to our fire behavior measurements via time stamps provided by the dataloggers. 
The Hettinger mesonet array is 3\textendash 8 km from burned pastures, while the Central Grasslands mesonet array is 1--7.5 km from burned pastures. 
We found a high degree of consistency between these meso-scale data and fire weather records made on the fireline during operational periods, and the open rangeland physiognomy with flat to rolling terrain precludes substantial microsite differences in weather between these records and the conditions at each fire behavior sample point.

\subsection{Data analysis}

@@

\textbf{Regression analysis.---}We assessed weather and fuel effects on three fire behavior response variables: Maximum flame temperature (\degC) 15 cm above the soil surface (mean of three thermocouples), maximum soil surface temperature (\degC; single thermocouple), and rate of spread (m min\textsuperscript{-1}) through each 1 m triangular microplot. 
Because all three response variables were best modeled with a gamma distribution, we fit generalized linear mixed-effect regression models for each response with the \texttt{glmer} function from the \emph{lme4} package in \textsf{R} \citep{bates2015}. 
Fixed effects consisted of weather and fuel variables, as described above. 
The random-effect term was constructed to account for random variance among
locations, spatial non-independence within locations and nested variance
within plots, and the effect of repeated measurements within each
location. 
Due to concerns about collinearity between relative humidity
and vapor pressure deficit because they are derived from the same
variables, vapor pressure deficit was excluded from regression analysis
for all three response variables.

\section{Results}

@@

Most measures of fuel, fire weather, and fire behavior showed considerable variability within each location, although rates of spread were generally low (Fig.~\ref{DataSummary}). 
Principal Components Analysis indicated fire behavior patterns were consistent across locations ($P$ = 0.11), although the fall season in Hettinger---our semi-arid location in southwestern North Dakota---tended to have drier air and hotter fires (Fig.~\ref{PCA}). 
Spring fires in the Central Grasslands were conducted under warmer and more evaporative (VPD) conditions than fall fires at Hettinger. 
The first two axes of the Principal Components Analysis (Fig.~\ref{PCA}) explained 86\% of overall variance in the fire behavior dataset. 
The first axis (52\% variance explained) was most strongly associated with flame temperature and rate of spread, while the second axis was more strongly associated with soil surface temperature. 
Dew point was marginally related ($P~<$ 0.05) and inversely associated with flame temperatures and rate of spread.

\begin{figure} 
\includegraphics[width = 1\columnwidth]{pca_gg-1.pdf}
\caption{Principal Components Analysis of fire behavior data (response
variables in blue; rate of spread (ROS), temperature above surface
(flame \degC), and temperature at soil surface (soil \degC) for prescribed
burns on rangeland at Hettinger (H), in southwestern North Dakota, and
Central Grasslands (CG), in central North Dakota. No difference between
locations ($P$ = 0.11). 
Total variance explained in these two axes = 86\%.}
\label{PCA} % (Fig.~\ref{PCA})
\end{figure}

@@

\section{Discussion}

@@

Mean temperatures recorded in this study are consistent with other
reports from northern rangelands. In central Alberta grassland,
\citet{bailey1980} observed that temperatures at the surface varied between
110\(^\circ\)C and 165\(^\circ\)C for backfires and headfires,
respectively, while at 15 cm above the ground, headfires averaged 200\(^\circ\)C; temperatures generally tracked with fuel load. 
Surface fires through jack pine barrens in Ontario had a similar range as ours
\citep[140-545\(^\circ\)C,][]{smith1966}. In our study, mean 15-cm
temperatures were 225\(^\circ\)C during spring burns in central North
Dakota and 250\(^\circ\)C during fall burns in southwestern North
Dakota; surface temperatures at both locations generally averaged just
above 100\(^\circ\)C (Fig.~\ref{DataSummary}).

Discrepancies between our data and others from the region are consistent with what would be expected when differences in the fire environment are considered. 
For example, ungrazed pastures in southwestern North Dakota with higher litter loads burned much hotter at the soil surface, at an average of 275$^\circ$C \citep{mcgranahan2022}.
\citet{ohrtman2015} reported a wide range of maximum temperatures at the soil surface---150-500\(^\circ\)C---that was generally explained by variability in annual productivity and clipping frequency, which altered fuel load. 
Our fires were also cooler than those reported by \citet{archibold2003} in Saskatchewan: using the mid-point of observations made at 10 cm and 20 cm as a comparison to our
15-cm values, spring fires reached 314\(^\circ\)C and fall fires reached
298\(^\circ\)C. But \citet{archibold2003} also reported substantially
lower fuel moisture in each season and they had approximately three
times the fuel load, likely due to an absence of grazing on the remnant
prairie. A previous study reported similar results---temperatures
approaching 500\(^\circ\)C when fuel loads averaged 2.8-4.5 t
ha\(^{\text{-}1}\) \citep{archibold1998}. 
With greater variability in fuel load and fuel moisture, we might also expect to see these factors have greater influence on aboveground flame temperatures. For example,
in Colorado, \citet{augustine2014} observed a strong linear relationship
between fuel load and temperatures ranging from 60-200\(^\circ\)C, but
their fuel load also ranged from 0.2 to 1.2 t ha\(^{\text{-}1}\).

Fires at both of our locations spread much more slowly than most reports
from other grasslands, due primarily to high fuelbed moisture content
and little opportunity for mitigation by wind or lower atmospheric
moisture (Fig.~\ref{DataSummary}).
Perhaps the most variability in fire spread was 
reported by \citet{sneeuwjagt1977} from prescribed grass fires in
California and Washington, where rates of spread ranged from 0.2-61 m
min\(^{\text{-}1}\). From grassland fires in South Africa and Kansas,
USA, \citet{trollope2002} reported average spread rates of 24 and 32 m
min\(^{\text{-}1}\), respectively, for head fires and 0.12 and 0.14 m
min\(^{\text{-}1}\), respectively, for back fires. In a tallgrass
prairie in Texas, \citet{clements2019} recorded fire spreading between
72-150 m min\(^{\text{-}1}\) with the wind and 48 m min\(^{\text{-}1}\)
for flanking fires. Likewise, fires through cured grass fuelbeds in
Australia spread much more rapidly than we observed---up to 18-180 m
min\(^{\text{-}1}\) \citep{cheney1993, cheney1995, cruz2015}. Through
partially-cured stands, though, \citet{cruz2015} found spread rates
dropped to 3-44 m min\(^{\text{-}1}\), approaching our location averages
of 2.1 and 3.1 m min\(^{\text{-}1}\). 
Consistent with our finding that only wind speed had a statistically-significant effect on increasing
rate of spread (Fig.~\ref{CIs}), \citet{cheney1993} found that wind was by far
the most important variable to spread rate.

@@

Much is made of the difference in fire behavior between head and back
fires in the fire ecology literature, and while we expect these
differences translate to different fire effects in our system, making
distinctions between fire types is difficult in both our data and our
management. \citet{trollope1978} emphasized that while head fires move
faster and generally release more energy, back fires effect greater
heating at the ground level. One would expect, then, that back fires
would have more opportunity to burn down through even thick \emph{P.
pratensis} litter to mineral soil. Unfortunately, our results offer
little insight into what fuel or weather variables enhance litter
combustion, likely because most of our fires never got very hot at the
soil surface  (Fig.~\ref{DataSummary}). 
Nor can we differentiate the direction of fire spread with the current trigonometry applied to the triangular thermocouple arrays \citep{simard1984}, although it would theoretically be possible to compare spread direction to wind direction if the latter data were available at a fine enough scale.
